---
title: "An illustration"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Required packages

The package ``gurobi'' requires installing Gurobi in local computer, and obtain a license from its website.

```{r}
require(Matrix)
require(gurobi)
```


## A function to generate optimal designs

This function generates optimal design with and without network connections. If no network adjacency matrix W is loaded in the function, the function will give optimal design with respect to covariates balance only. The default settings are $\alpha=0.001$, $\rho=0.5$,
and the timelimit to solve an optimization problem in Gurobi is 500 secs. 

```{r}
opt.design <- function(Z, W=NA, alpha=0.001, rho=0.5, timelimit=500) {
  ##################################################################
  # Function for optimal design with and without network
  # Inputs:
  # Z: covariate matrix, intercept not included
  # W: network adjacency matrix (NOT USED FOR network==F)
  # alpha: quantile parameter (NOT USED FOR network==F)
  # rho: network correlation parameter  (NOT USED FOR network==F)
  # timelimit (in sec): timelime in solving the optimization problem
  #
  # Outputs:
  # optimal allocation (-1, 1)
  #################################################################
  require(gurobi)
  require(Matrix)
  
  # common statistics
  Fmat <- cbind(1, Z)
  
  
  # setup global parameters
  params <- list(MIPGap=0.01, TimeLimit=timelimit)
  
  
  # setup optimization model
  model <- list()
  
  # setup linear constraint 
  model$A     <- matrix(1, 1, n)
  model$rhs   <- round(n/2) # to compile with the case that n is odd
  model$sense <- c("=")
  
  # set up decision variable type
  model$vtype <- rep('B', n)
  
  # with or without network
  
  network <- !is.na(sum(W))
  
  if(network == TRUE) {
    # network statistics
    n <- nrow(W)
    nedge <- rowSums(W)
    
    # set up quadratic objective
    Sigma.net <- Diagonal(n, x = nedge)-rho*W
    Fsig <- t(Fmat) %*% Sigma.net
    Sigma <- Fsig %*% Fmat		
    Q <- as.matrix(t(Fsig) %*% solve(Sigma, Fsig))
    model$Q     <- Q
    model$obj   <- -rowSums(Q)
    
    # set up quadratic constraint
    model$quadcon <- list()
    qc <- list()
    qc$Qc <- 4*W
    qc$q <- -4*rowSums(W)
    qc$rhs <- sqrt(sum(nedge))*qnorm(alpha)-sum(W)
    model$quadcon[[1]] <- qc
    
  } 

  if(network == FALSE) {
    # set up quadratic objective
    Finv <- solve(t(Fmat) %*% Fmat, t(Fmat))
    Q <- as.matrix(Fmat %*% Finv)
    model$Q     <- Q
    model$obj   <- -rowSums(Q)
  }
  
  # solve the problem with gurobi
  result <-gurobi(model, params)  
  return(2*result$x-1)
}
```


## Generate Data 

```{r}
n <- 100
p <- 10
net.p <- 0.08 # network density

# generate the covariates
Z <- matrix(sample(c(-1, 1), n*p, replace=TRUE), n, p)

# generate the network
W <- sample(c(0, 1), n*n, replace=TRUE, prob=c(1-net.p/2,net.p/2))
W <- matrix(W, n, n)
W <- 1*((W+t(W))>0)
diag(W) <- 0
```

## Generate Optimal Designs


```{r}
# optimal design with network 
x.opt <- opt.design(Z, W)
    
# optimal design without network
x.opt0 <- opt.design(Z)
```


## Evaluation Function

A function to compute the key statistics given a design. The value $\rho$ is the true correlation coefficient. The outputs are explained as follows:

- Impre: the improvement to precision in equation 17
- obj: the objective value in equation 9 of the  given design 
- Erand: expected objective value of the random design
- T1: T1 of the given design
- T2: T2 of the given design
- ET1: expected T1 value of the random design
- ET2: expected T2 value of the random design
- 1*(flag==0): a flag demonstrate an appropriate design is loaded. 

```{r}
evaluate <- function(x, W, Z, rho) {
  ##################################################################
  # Function to compute evaluation criterion for a design under car model assumption
  # Inputs:
  # x: design vector allocation (-1, 1)
  # W: network adjacency matrix 
  # Z: covariates
  # rho: network correlation parameter
  #
  # Outputs:
  # Precentage of Improvement to Precision and others
  #################################################################
  
  # compute statistics
  n <- nrow(W)
	nedge <- rowSums(W)
  nedge.total <- sum(W)
  p <- ncol(Z)
  Erand <- -999
  T1 <- -999
  T2 <- -999
  ImPre <- -999
  obj <- -999
  flag <- length(x)
  
  if(flag!=0) {
    Rmat <- Diagonal(n, x =nedge)-rho*W
    Fmat <- cbind(1, Z)
	  Fsig <- t(Fmat) %*% Rmat
	  xFsig <- Fsig %*% x
    Sigma <- Fsig %*% Fmat
          
    # compute the objective
    T1 <- rho * as.numeric(t(x)%*% W %*% x)
    T2 <- as.numeric(t(xFsig) %*% solve(Sigma, xFsig))
    obj <- nedge.total-T1-T2
          
    # compute the expected objective of random designs
    KF <- t(Fsig) %*% solve(Sigma, Fsig)
    K <- Rmat - KF
    Erand <- sum(diag(K))
    ET1 <- sum(diag(W))
    ET2 <- sum(diag(KF))
    
    # compute improvementment of precision 
    ImPre <- 1-Erand/obj
  }
  return(c(rho, ImPre, obj, Erand, T1, T2, ET1, ET2, 1*(flag==0)))
}
```


## Evaluation of the optimal designs

```{r}
rho_true <- 0.7
results <- evaluate(x.opt, W, Z, rho=rho_true)
results <- rbind(results, evaluate(x.opt0, W, Z, rho=rho_true))
results <- data.frame(results)
names(results) <- c("rho_true", "ImprovePrecision", "Obj", "E(obj)", "T1", "T2", "ET1", "ET2", "error")
results$network <- c(1, 0)
results
```


## Connection of submitted files to numerical sections


- Section 6: "synthetic.R" is the main file to generate all the results and figures.  

- Section 7: "real.R" is the main file to generate the results in Figure 7, and draw the figure; "robust_re_real.R" is the main file to generates the results in Figure 8, and "report_results_robust.R" generates Figure 8.

- Supplement: "gap.R" is the main file to generate figure S1. 



